%------------------------------------
% Dario Taraborelli
% Typesetting your academic CV in LaTeX
%
% URL: http://nitens.org/taraborelli/cvtex
% DISCLAIMER: This template is provided for free and without any guarantee
% that it will correctly compile on your system if you have a non-standard
% configuration.
% Some rights reserved: http://creativecommons.org/licenses/by-sa/3.0/
%------------------------------------

%!TEX TS-program = xelatex
%!TEX encoding = UTF-8 Unicode

\documentclass[11pt, a4paper]{article}
\usepackage{fontspec}


% DOCUMENT LAYOUT
\usepackage{geometry}
\geometry{a4paper, textwidth=5.5in, textheight=8.5in, marginparsep=7pt, marginparwidth=.6in}
\setlength\parindent{0in}

% FONTS
\usepackage{xunicode}
\usepackage{xltxtra}
\defaultfontfeatures{Mapping=tex-text} % converts LaTeX specials (``quotes'' --- dashes etc.) to unicode
\setromanfont [Ligatures={Common}, BoldFont={Gentium Basic Bold}, ItalicFont={Gentium Basic Italic}]{Gentium Basic}
\setmonofont[Scale=0.8]{Monaco}
% ---- CUSTOM AMPERSAND
\newcommand{\amper}{{\fontspec[Scale=.95]{Gentium Basic Italic}\selectfont\itshape\&}}
% ---- MARGIN YEARS
\usepackage{marginnote}
\newcommand{\years}[1]{\marginnote{\scriptsize #1}}
\renewcommand*{\raggedleftmarginnote}{}
\setlength{\marginparsep}{7pt}
\reversemarginpar

% HEADINGS
\usepackage{sectsty}
\usepackage[normalem]{ulem}
\sectionfont{\rmfamily\mdseries\upshape\Large}
\subsectionfont{\rmfamily\bfseries\upshape\normalsize}
\subsubsectionfont{\rmfamily\mdseries\upshape\normalsize}

% PDF SETUP
% ---- FILL IN HERE THE DOC TITLE AND AUTHOR
\usepackage[bookmarks, colorlinks, breaklinks, pdftitle={Robert Kubinec - vita},pdfauthor={Robert Kubinec}]{hyperref}
\hypersetup{linkcolor=black,citecolor=black,filecolor=black,urlcolor=black}
\usepackage{doi}

% DOCUMENT
\begin{document}
{\LARGE Robert Kubinec}\\[1cm]
University of South Carolina\\
Columbia, South Carolina\\[.2cm]
Email: \href{mailto:rkubinec@mailbox.sc.edu}{rkubinec@mailbox.sc.edu}\\
\textsc{url}: \href{http://www.robertkubinec.com}{www.robertkubinec.com}\\

%%\hrule
\section*{Current position}
\noindent
\years{2024-}\emph{Assistant Professor of Political Science}, Department of Political Science, University of South Carolina\\

\section*{Previous positions}
\noindent
\years{2019-2024}\emph{Assistant Professor of Political Science}, Social Sciences Division, New York University Abu Dhabi\\
\years{2018-2019}\emph{Postdoctoral Research Scholar}, Niehaus Center for Globalization and Governance, Princeton University\\

%\hrule
\section*{Education}
\noindent
\years{2018}\textsc{PhD} in Foreign Affairs, University of Virginia\\
\years{2010}\textsc{MA} in Middle East Studies, George Washington University\\
\years{2008}\textsc{BA} in International Relations, Wheaton College

\section*{Publications}

\subsection*{Books}

\years{2023} Kubinec, Robert. \emph{\href{https://www.cambridge.org/gb/academic/subjects/politics-international-relations/political-economy/making-democracy-safe-business-corporate-politics-during-arab-uprisings?format=HB}{Making Democracy Safe for Business: Corporate Politics During the Arab Uprisings}}. Cambridge University Press.
 
\subsection*{Journal Articles}

\years{2024} Kubinec, Robert; Carvalho, Luiz Max; Barcelo, Joan; Cheng, Cindy; Messerschmidt, Luca; Cottrell, Matthew Sean and Duba, Derek. ``A Bayesian Latent Variable Model for the Optimal Identification of Disease Incidence Rates Given Information Constraints." \emph{Journal of the Royal Statistical Society Series A}. Forthcoming. \doi{10.1093/jrsssa/qnae040}\\

\years{2024} Kubinec, Robert; Barcelo, Joan; Goldszmidt, Rafael; Grujic, Vanja; Model, Timothy; Schenk, Caress; Cheng, Cindy; Hale, Thomas; Spencer Hartnett; Allison, Messerschmidt, Luca; Petherick, Anna, and Thorvaldsdottir, Svanhildur. ``Cross-National Measures of the Intensity of COVID-19 Public Health Policies." Accepted at the \emph{Journal of Politics}. \doi{10.31235/osf.io/rn9xk}.\\

\years{2024} Cheng, Cindy; Messerschmidt, Luca; Bravo, Isaac; Waldbauer, Marco; Bhavikatti, Rohan; Schenck, Caress; Grujic, Vanja; Model, Tim, and Kubinec, Robert. ``Harmonizing Government Responses to the COVID-19 Pandemic." \emph{Scientific Data}. \doi{10.1038/s41597-023-02881-x}.\\

\years{2023} Cheng, Cindy; Messerschmidt, Luca; Bravo, Isaac; Waldbauer, Marco; Bhavikatti, Rohan; Schenck, Caress; Grujic, Vanja; Model, Tim, and Kubinec, Robert. ``A General Primer for Harmonizing Data." \emph{Scientific Data}. \doi{10.1038/s41597-024-02956-3j}.\\

\years{2023} Kubinec, Robert and Milner, Helen. ``Taxes in the Time of Revolution: An Experimental Test of the Rentier State during Algeria's Hirak." \emph{World Politics}. \doi{10.1353/wp.2024.a924508}.\\

\years{2023} Kubinec, Robert; Lee, Haillie and Tomashevskiy, Andrey. ``Why Corporate Political Connections Can Impede Investment." \emph{Comparative Political Studies} \doi{10.1177/00104140231204227}.\\

\years{2022} Kilavuz, Tahir; Grewal, Sharan and Kubinec, Robert. ``Ghosts of the Black Decade: How Legacies of Violence Shaped Algeria's Hirak Protests." \emph{Journal of Peace Research} \doi{10.1177/00223433221137613}.\\

\years{2022}  Kubinec, Robert.``Ordered Beta Regression: A Parsimonious, Well-Fitting Model for Continuous Data with Lower and Upper Bounds." \textit{Political Analysis} \doi{10.1017/pan.2022.20}.\\

\years{2022}Barcelo, Joan; Kubinec, Robert; Cheng, Cindy; Hoye Rahn, Tiril; and Messerschmidt, Luca. ``Windows of Repression: Using COVID-19 Policies against Political Dissidents?"  \textit{Journal of Peace Research} \doi{10.1177/00223433211062389}.\\

\years{2021}Kubinec, Robert; Lee, Haillie and Tomashevskiy, Andrey. ``Politically-Connected Companies Are Less Likely to Shutdown Due to COVID-19 Restrictions." \textit{Social Science Quarterly}. \doi{10.1111/ssqu.13040}.\\

\years{2020}Kubinec, Robert and Owen, John. ``When Groups Fall Apart: Identifying Transnational Polarization with Twitter from the Arab Uprisings." \textit{Political Analysis} \doi{10.1017/pan.2020.46}. \\

\years{2020}Cheng, Cindy; Barcelo, Joan; Hartnett, Allison; Kubinec, Robert and Messerschmidt, Luca. ``COVID-19 Government Response Event Dataset (CoronaNet v.1.0)." \textit{Nature Human Behavior}. \doi{10.1038/s41562-020-0909-7}.\\

\years{2020}Kropko, Jonathan and Kubinec, Robert. ``Why the Two-way Fixed Effects Model is Difficult to Interpret, and What to Do About It." \textit{PLOS One}. \doi{10.1371/journal.pone.0231349}

\end{document}
